The Minimal Retroreflective Microfacet Model

### 2025-07-23 article.tex
further cleanup for refraction, fix some formulas and notation

\begin{figure}
  \begin{center}
    \begin{tabular} {|p{0.2\textwidth}|p{0.7\textwidth}|}
%    \begin{tabular} {|l|l|}
      \hline
      $\vecv$ & normalized vector to viewer \\
      \hline
      $\vecl$ & normalized vector to light \\
      \hline
      $\vecn$ & surface normal \\
      \hline
      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the normal) \\
      \hline
      $\vechr$ & half vector of reflection, $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
      \hline
      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to \\
      \hline
      $\vdot{\vecx}{\vecy}$ & dot product of $\vecx$ and $\vecy$ \\
      \hline
      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e. $\text{reflect}(\mathbf{x}, \mathbf{m}) = -\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
      \hline
      $\Omega$ & hemisphere of directions around the normal \todo{need more formalization for transmission}\\
      \hline
    \end{tabular}
  \end{center}
  \caption{Notations used.}
  \label{fig:notation}
\end{figure}

Using the notations from Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
\begin{equation}
  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl) \, F(\vechr, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
  \label{eq:microfacet_brdf}
\end{equation}
where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vechr) \vdot{\vecn}{\vechr} d\vechr = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term. The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
Similarly, using the half-vector of refraction as selector, yields the corresponding BTDF
\begin{equation}
  f_t(\vecv, \vecl) =
  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\etav^2 D(\vecht) \, G_{2}(\vecv,\vecl) \, \left(1 - F(\vecht, \vecv) \right)}
       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \avdot{\vecv}{\vecht} + \etal \avdot{\vecl}{\vecht} \right)^2}
  \label{eq:microfacet_btdf}.
\end{equation}
In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
In this work, we show that we can use the back vector~\cite{BelcourBackvector} used to model retro-reflection as a microfacet selector. Before detailing the impact of such a change, we review some properties of microfacet models.

\paragraph*{Shadowing-Masking.}
The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vecl)$ for view direction and
$G_1(\vecl, \vecv)$ for light direction:
\begin{equation}
G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right)
\end{equation}
Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
$G_{2}(\vecv, \vecl) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
\begin{equation}\label{eq:visible}
  \vdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
\end{equation}
Choices for $G_1$ with this property are the Smith masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
\end{equation}
where analytic expressions are available for Beckmann and GGX distributions, and the v-cavity masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
\end{equation}
which is generally applicable for distributions with symmetry around the normal.

@@

As illustrated in figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
While there is no immediate physical foundation to this construction, this interpretation can serve as motivation why components from the regular microfacet model still work when changing the domain to the back-vector.

\paragraph*{Jacobian of Back-Vector.}
First we observe that  Jacobian of the change of variable from the back-vector to the light vector $\vecl$~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|}
\end{equation}
\todo{Formulate this for refraction?}

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead. It follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb, \vecm) }{4 \left| \vdot{\vecvv}{\vecb} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
\todo{Formulate this for refraction?}

@@

\paragraph*{Final form.}
The final retro-reflective microfacet model can be written as
\begin{equation}
  %  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{retro}(\vecv, \vecl) = f_{r, \text{retro}}(\vecv, \vecl) + f_{t, \text{retro}}(\vecv, \vecl) =
  f_r(\vecvv, \vecl) + f_t(\vecvv, \vecl)
  \label{eq:retro_microfacet_model}.
\end{equation}
\todo{separte BRDF and BTDF}
In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
Note that for this we impose one technical assumption on $D$:
we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).

@@

In analogy to $\vecvv$ we let $\vecll := \text{reflect}(\vecl, \vecn)$.

@@

As $D$ is reflection-symmetric around the normal, i.e. it takes the form $D(\vech) = D\left(\vdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with a back-vector obeys symmetry.
Further, the reflection symmetry of $D$ along with equation \eqref{eq:nk} gives

@@

  \frac{\vdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech} \\
  &= \frac{\vdot{\vecvv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecvv}{\vech} d\vech} =

@@

  \frac{1 + \vdot{\vecyy}{\vecxx}}{\lVert \vecyy + \vecxx \rVert} =

@@

  \frac{-\etax \vdot{\vecxx}{\vecxx} - \etay\vdot{\vecxx}{\vecyy}}{\lVert -\etax\vecxx  -\etay\vecyy \rVert} =
  \vdot{\vecxx}{\vech(\vecyy,\vecxx)}.

@@

Further, using equations \eqref{eq:khr} and \eqref{eq:kht} along with equations \eqref{eq:nh} and \eqref{eq:nk} we get

@@

In summary, we can conclude the symmetry of both the BRDF

@@

  f_{r,\text{retro}}(\vecl, \vecv)

@@

and the BTDF

@@

    {\avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \avdot{\vecvv}{\vecht} + \etal \avdot{\vecl}{\vecht}\right)^2} \\

@@

    {\avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} \left(\etav \avdot{\vecvv}{\vecht} + \etal \avdot{\vecl}{\vecht}\right)^2} \\

@@

  \rho_{\text{retro}}(\vecv) =
  \int_\Omega f_{\text{retro}}(\vecv, \vecl) \vdot{\vecl}{\vecn} d\vecl =
  \int_\Omega f(\vecvv, \vecl) \vdot{\vecl}{\vecn} d\vecl = \rho(\vecvv)

@@

i.e. we get the albedo of the regular microfacet BRDF for the reflected direction.
As the original microfacet BRDF fulfills energy conversion, the retro-reflective modification does, too.

@@

\begin{itemize}
\item
  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
\item
  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BSDF.
  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
\item
  As the albedos of standard BSDF and retro-reflective BSDF are essentially identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
\item Pseudo-code is provided in Listing~\ref{lst:implementation}.
  \todo{mraab: I think source code may be overkill, given the simplicity of the model.}
\end{itemize}

\noindent\begin{minipage}[tb]{\linewidth}
\begin{lstlisting}[
    language=GLSL,
    caption=Pseudo code to implement the retro-reflective microfacet model.,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
vec3 brdf_retro(vec3 L, vec3 V, vec3 N, vec3 X, vec3 Y)
{
    vec3 Vp = -reflect(V,N); // jp - is this wrong? Why the minus sign

    // The following could all be replaced with:
    // We could give the simplified form of brdf_ggx, but not essential.
    return brdf_ggx(L, Vp, N, X, Y);

    /*
    float NdotL = dot(N, L);
    float NdotV = dot(N, Vp);
    if (NdotL < 0 || NdotV < 0) return vec3(0);

    vec3 B = normalize(L+Vp);
    float NdotH = dot(N, B);
    vec3 wB = vec3(dot(B,X),  dot(B,Y),  dot(B,N));
    vec3 wL = vec3(dot(L,X),  dot(L,Y),  dot(L,N));
    vec3 wV = vec3(dot(Vp,X), dot(Vp,Y), dot(Vp,N));

    float D = D_GGX(wB, ax, ay);
    float G = G2_GGX(wV, wL, ax, ay);

    float c = dot(L, B);
    float F = schlick(f0, c); // I don't think we need to give the form of the Fresnel

    float val = Ks * D * G * F / (4 * NdotL * NdotV);
    return vec3(val);
    */
}
\end{lstlisting}
\end{minipage}

>>>>

\begin{figure}
  \begin{center}
    \begin{tabular} {|p{0.2\textwidth}|p{0.7\textwidth}|}
%    \begin{tabular} {|l|l|}
      \hline
      $\vecv$ & normalized vector to viewer \\
      \hline
      $\vecl$ & normalized vector to light \\
      \hline
      $\vecn$ & surface normal, with the convention that it points to
      the medium with lower index of refraction in cases dealing with $\vecv$ and $\vecl$ pointing to different sides\\
      \hline
      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the normal) \\
      \hline
      $\vechr$ & half vector of reflection, $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
      \hline
      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to (always points to medium with lower index of refraction)\\
      \hline
      $\vdot{\vecx}{\vecy}$ & dot product of $\vecx$ and $\vecy$ \\
      \hline
      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e. $\text{reflect}(\mathbf{x}, \mathbf{m}) = -\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
%      \hline
%      $\Omega$ & hemisphere of directions around the normal\\
      \hline
      $\Omega, \Omega_{\vecx}$ & hemisphere of directions around the normal, hemisphere of directions around the normal pointing to the side $\vecx$ points to\\
      \hline
    \end{tabular}
  \end{center}
  \caption{Notations used.}
  \label{fig:notation}
\end{figure}

Using the notations from Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
\begin{equation}
  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl) \, F(\vechr, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
  \label{eq:microfacet_brdf}
\end{equation}
where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term. The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
Similarly, using the half-vector of refraction as selector, yields the corresponding BTDF
\begin{equation}
  f_t(\vecv, \vecl) =
  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\etav^2 D(\vecht) \, G_{2}(\vecv,\vecl) \, \left(1 - F(\vecht, \vecv) \right)}
       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecv}{\vecht} + \etal \vdot{\vecl}{\vecht} \right)^2}
  \label{eq:microfacet_btdf}.
\end{equation}
In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
In this work, we show that we can use the back vector~\cite{BelcourBackvector} used to model retro-reflection as a microfacet selector. Before detailing the impact of such a change, we review some properties of microfacet models.

\paragraph*{Shadowing-Masking.}
The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vecl)$ for view direction and
$G_1(\vecl, \vecv)$ for light direction:
\begin{equation}
G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right)
\end{equation}
Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
$G_{2}(\vecv, \vecl) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
\begin{equation}\label{eq:visible}
  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
\end{equation}
Choices for $G_1$ with this property are the Smith masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
\end{equation}
where analytic expressions are available for Beckmann and GGX distributions, and the v-cavity masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
\end{equation}
which is generally applicable for distributions with symmetry around the normal.

@@

As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
While there is no immediate physical foundation to this construction, this interpretation can serve as motivation why components from the regular microfacet model still work when changing the domain to the back-vector.

\paragraph*{Jacobian of Back-Vector.}
First we observe that  Jacobian of the change of variable from the back-vector to the light vector $\vecl$~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|}
\end{equation}
\todo{Formulate this for refraction or remove/simplify?}

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead. It follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb, \vecm) }{4 \left| \vdot{\vecvv}{\vecb} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
\todo{Formulate this for refraction or remove/simplify?}

@@

\paragraph*{Final form.}
The final retro-reflective microfacet model can be written as
\begin{equation} \label{eq:retro_microfacet_model}
  %  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{retro}(\vecv, \vecl) =
  \begin{cases}
    f_{r, \text{retro}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if} \vdot{\vecv}{\vecn}\vdot{\vecl}{\vecn} \geq 0, \\
    f_{t, \text{retro}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
  \end{cases}
\end{equation}
In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
Note that for this we impose one technical assumption on $D$:
we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).

@@

In analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.

@@

As $D$ is reflection-symmetric around the normal, i.e. it takes the form $D(\vech) = D\left(\avdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with a back-vector obeys symmetry.
Further, the reflection symmetry of $D$ along with Equation \eqref{eq:nk} gives

@@

  \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech} \\
  &= \frac{\avdot{\vecvv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecvv}: \vdot{\vecvv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecvv}{\vech} d\vech} =

@@

  \frac{\vdot{\vecyy}{\vecyy} + \vdot{\vecyy}{\vecxx}}{\lVert \vecyy + \vecxx \rVert} =

@@

  \frac{-\etax \vdot{\vecyy}{\vecyy} - \etay\vdot{\vecxx}{\vecyy}}{\lVert -\etax\vecxx  -\etay\vecyy \rVert} =
  \vdot{\vecyy}{\vech(\vecyy,\vecxx)}.

@@

Further, using Equations \eqref{eq:khr} and \eqref{eq:kht} along with Equations \eqref{eq:nh} and \eqref{eq:nk} und using $\vech(\vecx,\vecy) = \vech(\vecy,\vecx)$ we get

@@

In summary, we can conclude the symmetry of the BRDF

@@

  f_{r,\text{retro}}(\vecl, \vecv).

@@

Analogously, we obtain the symmetry requirement of the BTDF

@@

    {\avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecvv}{\vecht(\vecvv, \vecl)} + \etal \vdot{\vecl}{\vecht(\vecvv, \vecl)}\right)^2} \\

@@

    {\avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} \left(\etav \vdot{\vecv}{\vecht(\vecv, \vecll)} + \etal \vdot{\vecll}{\vecht(\vecv, \vecll)}\right)^2} \\

@@

  \rho_{r, \text{retro}}(\vecv) =
  \int_{\Omega_{\vecv}} f_{r, \text{retro}}(\vecv, \vecl) \vdot{\vecl}{\vecn} d\vecl =
  \int_{\Omega_{\vecvv}} f_r(\vecvv, \vecl) \vdot{\vecl}{\vecn} d\vecl = \rho_r(\vecvv),\\
  \rho_{t, \text{retro}}(\vecv) =
  \int_{\Omega_{-\vecv}} f_{t, \text{retro}}(\vecv, \vecl) \avdot{\vecl}{\vecn} d\vecl =
  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} d\vecl = \rho_t(\vecvv),

@@

i.e. we get the albedo of the regular microfacet BSDF for the reflected direction.
As the original microfacet BSDF fulfills energy conversion, the retro-reflective modification does, too.

@@

\begin{itemize}
\item
  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
\item
  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BSDF.
  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
\item
  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
\item Pseudo-code is provided in Listing~\ref{lst:implementation}.
\end{itemize}

\noindent\begin{minipage}[tb]{\linewidth}
\begin{lstlisting}[
    language=GLSL,
    caption=Pseudo code to implement the retro-reflective microfacet model using existing regular microfacet BSDF implementation,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return sample_bsdf(V, L, N, params);
}
Eval_result eval_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return eval_bsdf(V, L, N, params);
}

\end{lstlisting}
\end{minipage}
